Spatial Analysis

### 2021-07-08 Scripts/spatial_analisis.Rmd
First try to represent offset

@@

```{r paper_figure_area_map, eval = F, echo = F, results='hide'}

@@

### Aggregated Proportional change

@@

land_sf %>% 
  left_join(state_fit,
            by = "state") %>% 
  ggplot() +
  geom_sf(aes(fill = mean_per)) +
  viridis::scale_fill_viridis("Average proportion per State", alpha = 0.8) +
  facet_wrap(~ label, nrow = 1) +
  labs(x = "Longitude", 
       y = "Latitude") +
  my_ggtheme_p(facet_tx_s = 16,
               leg_pos = "bottom",
               leg_tl_s = 14,
               ax_tx_s = 10,
               ax_tl_s = 14) +
  theme(legend.key.width = unit(2,"line")) +
  ggsave(filename = "proportion_chg_agg.jpg",
         plot = ggplot2::last_plot(),
         path = my_path("R","Figures"),
         width = 9,
         height = 5
         )

@@

p <- land_sf %>% 
  left_join(state_fit,
            by = "state") %>% 
  ggplot() +
  # geom_sf(aes(fill = mean_per)) +
  geom_sf(aes(fill = mean_per, text = paste(state, mean_per,"% of proportion"))) +
  viridis::scale_fill_viridis("Average proportion per State", alpha = 0.8) +
  facet_wrap(~ region + label, 
             nrow = 2) +
  labs(x = "Latitude", 
       y = "Longitude") +
  my_ggtheme_p(facet_tx_s = 12,
               leg_pos = "bottom") +
  theme(legend.key.width = unit(1,"line"))

@@

# Plot me
p <- ggplot(state_fit) +
  geom_area(
    aes(
      x = year,
      y = round(percentage),
      fill = state
    )
  ) +
  ylab("Percentage (%)") +
  # viridis::scale_fill_viridis(discrete = T, alpha = 0.8) +
  scale_fill_manual(values = state_pallet) +
  MyFunctions::my_ggtheme_p() +
  facet_wrap(~region, ncol = 1)

@@

# Plot me
p <- ggplot(state_fit) +
  geom_area(
    aes(
      x = year,
      y = round(RMean),
      fill = state
      # fill = order
    )
  ) +
  ylab("10 yrs running average (%)") +
  scale_fill_manual(
    "State",
    values = state_pallet,
    # labels = state_fit %>% arrange(order) %>%  pull(state) %>% unique()
  ) +
  MyFunctions::my_ggtheme_p() +
  facet_wrap(~region, ncol = 1)

@@

# Plot me
ggplot(state_fit) +
  geom_area(
    aes(
      x = year,
      y = round(RMean),
      fill = order
      # fill = state
    )
  ) +
  ylab("Propotion (10 yrs running average)") +
  # viridis::scale_fill_viridis(discrete = T, alpha = 0.8) +
  scale_fill_manual(
    "State",
    values = state_pallet,
    labels = state_fit %>% arrange(order) %>%  pull(state) %>% unique()
  ) +
  my_ggtheme_p(leg_pos = "right") +
  theme(legend.key.width = unit(1,"line")) + 
  ggsave(filename = "area_plot_avg_agg.jpg",
         plot = ggplot2::last_plot(),
         path = my_path("R","Figures"),
         width = 8,
         height = 6
  )

@@

### Per Species Proportional change

@@

# Plot me
ggplot(state_fit) +
  geom_area(
    aes(
      x = year,
      y = round(RMean),
      fill = order
      # fill = state
    )
  ) +
  ylab("Propotion (10 yrs running average)") +
  # viridis::scale_fill_viridis(discrete = T, alpha = 0.8) +
  scale_fill_manual(
    "State",
    values = state_pallet,
    labels = state_fit %>% arrange(order) %>%  pull(state) %>% unique()
  ) +
  my_ggtheme_p(leg_pos = "right") +
  theme(legend.key.width = unit(1,"line")) + 
  facet_wrap(~spp, ncol = 1, strip.position = "right") +
  ggsave(filename = "area_plot_avg_spp.jpg",
         plot = ggplot2::last_plot(),
         path = my_path("R","Figures"),
         width = 10,
         height = 12
  )

@@

### Prepare data

@@

### Difference line plot

@@

quota_df %>%
  mutate(diff = Historic- Distribution,
         diff_rmean = RMean- Distribution) %>% 
  gather("type","value",diff:diff_rmean) %>% 
  # View()
  ggplot() +
  geom_line(
    aes(
      x = year,
      y = value,
      color = order
    )
  ) +
  scale_color_manual("State",
    values = state_pallet,
    labels = quota_df %>% arrange(order) %>%  pull(state) %>% unique()
    ) +
  facet_grid(type~spp, 
             scales = "free")

```
### Linetype plot

@@

quota_df %>% 
  gather("level","quota",Distribution:RMean) %>% 
  # mutate(diff = quota_per- percentage) %>% 
  # View()
  ggplot() +
  geom_line(
    aes(
      x = year,
      y = quota,
      color = order
    )
  ) +
  labs(x = "Year",y = "Quota (%)") +
  scale_fill_manual("State",
    values = state_pallet,
    labels = quota_df %>% arrange(order) %>%  pull(state) %>% unique()
    ) +
  scale_color_manual("State",
    values = state_pallet,
    labels = quota_df %>% arrange(order) %>%  pull(state) %>% unique()
    ) +
  scale_linetype("Quota") +
  facet_grid(spp ~ level)

```
### Efficiency index

@@

$$ei = \frac{HistoricQuota}{DistributionProportion}$$ 
So, in cases qhere *ei* = 1, the quota is aligned with the distribution; when *ei* > 1 then the historic quota is more than the stock's State's distribution, contrarily, *ei* < 1 means that the historic quota is less than what the State currently has.

**Note:** There are some crazzy outliers that have been removed, for now...

@@

quota_df %>% 
  # filter(
  #   spp =="Centropristis striata"
  # ) %>% 
  mutate(index = Distribution/Historic,
         index_rmean = RMean/Historic) %>%
  gather("level","index",index:index_rmean) %>%
  filter(index < 5) %>% # there are some craaaazy outliers
  # View()
  ggplot() +
  geom_line(
    aes(
      x = year,
      y = index,
      color = state
    ),
  ) +
  labs(x = "Year",y = "Efficiency index") +
  scale_fill_manual("State",
    values = state_pallet,
    labels = quota_df %>% arrange(order) %>%  pull(state) %>% unique()
    ) +
  scale_color_manual("State",
    values = state_pallet,
    labels = quota_df %>% arrange(order) %>%  pull(state) %>% unique()
    ) +
  facet_grid(level~spp,
             scales = "free")

```
### Efficiency index

>>>>

### Dynamic figure: Aggregated Proportional change per survey

p <- land_sf %>% 
  left_join(state_fit,
            by = "state") %>% 
  ggplot() +
  # geom_sf(aes(fill = mean_per)) +
  geom_sf(aes(fill = mean_per, text = paste(state, mean_per,"% of proportion"))) +
  viridis::scale_fill_viridis("Average proportion per State", alpha = 0.8) +
  facet_wrap(~ region + label, 
             nrow = 2) +
  labs(x = "Latitude", 
       y = "Longitude") +
  my_ggtheme_p(facet_tx_s = 12,
               leg_pos = "bottom") +
  theme(legend.key.width = unit(1,"line")) +
  theme_dark()

@@

```{r paper_figure_area_map, eval = F, echo = T, results='hide'}

@@

### Paper figure: Aggregated Proportional change

@@

agg_map <- land_sf %>% 
  left_join(state_fit,
            by = "state") %>% 
  ggplot() +
  geom_sf(aes(fill = mean_per)) +
  viridis::scale_fill_viridis("Average proportion per State", alpha = 0.8) +
  facet_wrap(~ label, nrow = 1) +
  labs(x = "Longitude", 
       y = "Latitude") +
  my_ggtheme_p(facet_tx_s = 16,
               leg_pos = "bottom",
               leg_tl_s = 14,
               ax_tx_s = 10,
               ax_tl_s = 14) +
  theme(legend.key.width = unit(2,"line"))

  ggsave(filename = "proportion_chg_agg.jpg",
         plot = agg_map,
         path = my_path("R","Figures"),
         width = 9,
         height = 5
         )

@@

![Proportion change aggregated all species map for paper](/Volumes/Enterprise/Data/AcrossBoundaries/Results/Figures/proportion_chg_agg.jpg)

@@

# Plot me
p <- ggplot(state_fit) +
  geom_area(
    aes(
      x = year,
      y = round(percentage),
      fill = state
    )
  ) +
  ylab("Percentage (%)") +
  # viridis::scale_fill_viridis(discrete = T, alpha = 0.8) +
  scale_fill_manual(values = state_pallet) +
  MyFunctions::my_ggtheme_p() +
  facet_wrap(~region, ncol = 1) +
  theme_dark()

@@

# Plot me
p <- ggplot(state_fit) +
  geom_area(
    aes(
      x = year,
      y = round(RMean),
      fill = state
      # fill = order
    )
  ) +
  ylab("10 yrs running average (%)") +
  scale_fill_manual(
    "State",
    values = state_pallet,
    # labels = state_fit %>% arrange(order) %>%  pull(state) %>% unique()
  ) +
  MyFunctions::my_ggtheme_p() +
  facet_wrap(~region, ncol = 1) +
  theme_dark()

@@

# Plot me
agg_area <- ggplot(state_fit) +
  geom_area(
    aes(
      x = year,
      y = round(RMean),
      fill = order
      # fill = state
    )
  ) +
  ylab("Propotion (10 yrs running average)") +
  # viridis::scale_fill_viridis(discrete = T, alpha = 0.8) +
  scale_fill_manual(
    "State",
    values = state_pallet,
    labels = state_fit %>% arrange(order) %>%  pull(state) %>% unique()
  ) +
  my_ggtheme_p(leg_pos = "right") +
  theme(legend.key.width = unit(1,"line"))

  ggsave(filename = "area_plot_avg_agg.jpg",
         plot = agg_area,
         path = my_path("R","Figures"),
         width = 8,
         height = 6
  )

@@

![Proportion change aggregated all species map for paper](/Volumes/Enterprise/Data/AcrossBoundaries/Results/Figures/area_plot_avg_agg.jpg)

### Paper figure: Per Species Proportional change

@@

# Plot me
ssp_area <- ggplot(state_fit) +
  geom_area(
    aes(
      x = year,
      y = round(RMean),
      fill = order
      # fill = state
    )
  ) +
  ylab("Propotion (10 yrs running average)") +
  # viridis::scale_fill_viridis(discrete = T, alpha = 0.8) +
  scale_fill_manual(
    "State",
    values = state_pallet,
    labels = state_fit %>% arrange(order) %>%  pull(state) %>% unique()
  ) +
  my_ggtheme_p(leg_pos = "bottom") +
  theme(legend.key.width = unit(1,"line")) + 
  facet_wrap(~spp, ncol = 1, strip.position = "right")

  ggsave(filename = "area_plot_avg_spp.jpg",
         plot = ssp_area,
         path = my_path("R","Figures"),
         width = 10,
         height = 12
  )

```
![Proportion change aggregated all species map for paper](/Volumes/Enterprise/Data/AcrossBoundaries/Results/Figures/area_plot_avg_spp.jpg)

@@

### Get data ready

@@

### Exploring different ways to show the results

#### Difference line plot

@@

quota_df %>%
  mutate(diff = Historic- Distribution,
         diff_rmean = RMean- Distribution) %>% 
  gather("type","value",diff:diff_rmean) %>% 
  # View()
  ggplot() +
  geom_line(
    aes(
      x = year,
      y = value,
      color = order
    )
  ) +
  scale_color_manual("State",
    values = state_pallet,
    labels = quota_df %>% arrange(order) %>%  pull(state) %>% unique()
    ) +
  facet_grid(type~spp, 
             scales = "free") +
  theme_dark()

```
#### Linetype plot

@@

quota_df %>% 
  gather("level","quota",Distribution:RMean) %>% 
  # mutate(diff = quota_per- percentage) %>% 
  # View()
  ggplot() +
  geom_line(
    aes(
      x = year,
      y = quota,
      color = order
    )
  ) +
  labs(x = "Year",y = "Quota (%)") +
  scale_fill_manual("State",
    values = state_pallet,
    labels = quota_df %>% arrange(order) %>%  pull(state) %>% unique()
    ) +
  scale_color_manual("State",
    values = state_pallet,
    labels = quota_df %>% arrange(order) %>%  pull(state) %>% unique()
    ) +
  scale_linetype("Quota") +
  facet_grid(spp ~ level)+
  theme_dark()

```
#### Efficiency index

@@

$$ei = \frac{HistoricQuota}{DistributionProportion}$$ 
So, in cases where *ei* = 1, the quota is aligned with the distribution; when *ei* > 1 then the historic quota is more than the stock's State's distribution, contrarily, *ei* < 1 means that the historic quota is less than what the State currently has.

**Note:** 
- There are some crazzy outliers that have been removed, for now... 
- Bottom plot representing index as a Rmean

##### As a normal year to year

@@

eff_index <- quota_df %>% 
  # filter(
  #   spp =="Centropristis striata"
  # ) %>% 
  mutate(index = Distribution/Historic,
         index_rmean = RMean/Historic) %>%
  gather("level","index",index:index_rmean) %>%
  filter(index < 5) # there are some craaaazy outliers
  # View()

ggplot(eff_index) +
  geom_line(
    aes(
      x = year,
      y = index,
      color = state
    ),
  ) +
  labs(x = "Year",y = "Efficiency index") +
  scale_fill_manual("State",
    values = state_pallet,
    labels = quota_df %>% arrange(order) %>%  pull(state) %>% unique()
    ) +
  scale_color_manual("State",
    values = state_pallet,
    labels = quota_df %>% arrange(order) %>%  pull(state) %>% unique()
    ) +
  facet_grid(level~spp,
             scales = "free")+
  theme_dark()
